\section*{Appendix B: Tensor Notation} \label{app:AppendixB}

Tensors of various ranks are mathematical entities are will be encountered with great frequency in all or your core courses. In physical science, tensors characterize the properties of a physical system. Tensors are the de facto tool used to describe, for example, diffusion, nucleation and growth, states of stress and strain, Hamiltonians in quantum mechanics, and many, many, more physical phenomenon. This is because physical processes of interest to Materials Scientists take place in Eucledian 3-space ($\varmathbb{R}^3$).

The following brief introduction has been adapted from \href{https://www.grc.nasa.gov/www/k-12/Numbers/Math/documents/Tensors_TM2002211716.pdf}{\textit{An Introduction to Tensors for Students of Physics and Engineering}}. We build up our description of the notation and handling of tensors starting with by separately describing rank-0, rank-1, rank-2, and rank-3 tensors. These are referred to respectively as scalars, vectors, matrices, and tensors (rank $\geq3$). \emph{Tensors} are a generalization of vectors and matrices that you have encountered into higher dimensionality. Classifications of tensors by order or rank allows us to quickly identify the number of tensor components we will work with: a tensor of order $p$ has $N^p$ components, where $N$ is the dimentionality of space in which we are operating. In general, you will be operating in Eucledian 3-space, so the numer of components of a tensor is defined as $3^p$

We will address this by considering the question: ``how do I get to your apartment'' under a variety of circumstances. 

\paragraph{Scalars} are considered tensors of \emph{order} or \emph{rank} of zero. Scalars represent physical quantities (often accompanied by a unit of measurement) that possess only a magnitude: e.g., temperature, mass, charge, and distance. Scalars are typically represented by Latin or Greek symbols and are have $3^{0} = 1$ component.

\paragraph{Vectors} are tensors of \emph{rank} of 1. Scalars represent physical quantities (often accompanied by a unit of measurement) that possess only a magnitude: e.g., temperature, mass, charge, and distance. Vectors are typically represented using lowercase bold, italic (or sometimes not) symbols such as $v \text{or ,} a$. In $\varmathbb{R}^3$ we have and are have $3^{1} = 3$ components.
